---
title: "Comparative functional enrichment analysis between proteins annotated as displaying intensity variation in one of the four ciliary locations and annotated as stable"
author: "Konstantin Kahnert"
date: "`r Sys.Date()`"
output:
  pdf_document:  
    latex_engine: xelatex   
    keep_tex: false
  html_document:
    toc: true
    df_print: paged
    toc_float: true
    number_sections: true
    theme: flatly
    highlight: tango
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = "cairo_pdf", message = FALSE, warning = FALSE)
```

# Load libraries

```{r}
library(tidyverse)          
library(clusterProfiler)    
library(enrichplot)         
library(org.Hs.eg.db)  
library(Cairo) 
library(svglite)
library(VennDiagram)       
library(grid)               
library(simplifyEnrichment)
library(readxl)
```

# Define functions

```{r}
# map_cluster_number: 
#   - x: enrichResult or compareClusterResult (from clusterProfiler)
#   - df: simplifyGO result data.frame with columns ID (term IDs) and Cluster_num
#   - comp: if TRUE, use x@compareClusterResult, else x@result
# Returns a data.frame of all enrich columns + term_size + Cluster_num
map_cluster_number <- function(x,
                               df,
                               comp = FALSE) {
  ## 1. Standardize the input DF’s column names
  colnames(df) <- c("ID", "Cluster_num")

  ## 2. Select the correct slot from the clusterProfiler object
  if (comp) {
    # from compareClusterResult
    results <- x@compareClusterResult
  } else {
    # from a single enrichResult
    results <- x@result
  }

  ## 3. Get term_size (the numerator of the "BgRatio" string)
  ##    e.g. "12/21273" → 12
  results$term_size <- as.numeric(
    sapply(strsplit(results$BgRatio, "/"),
           function(parts) as.numeric(parts[1]))
  )

  ## 4. Merge the enrichment results with the cluster assignments
  merged_results <- merge(
    results,
    df,
    by    = "ID",
    all.x = TRUE
  )

  ## 5. Return the processed data frame
  return(merged_results)
}


# Function to map Ensembl IDs to gene symbols for a single row
map_geneID_to_symbol <- function(geneID_str) {
  geneIDs <- unlist(strsplit(geneID_str, "/"))
  mapped_ids <- bitr(geneIDs, 
                     fromType = "ENSEMBL", 
                     toType = "SYMBOL", 
                     OrgDb = org.Hs.eg.db)
  
  # drop duplicates in ensembl column and keep first occurence
  mapped_ids <- mapped_ids[!duplicated(mapped_ids$ENSEMBL),]
  
  gene_symbols <- mapped_ids$SYMBOL
  return(paste(gene_symbols, collapse = "/"))
}


# Function to map Ensembl IDs to gene symbols for a single row
map_geneID_to_name <- function(geneIDs) {
  mapped_ids <- bitr(geneIDs, 
                     fromType = "ENSEMBL", 
                     toType = c("GENENAME", "SYMBOL"), 
                     OrgDb = org.Hs.eg.db)
  
  # drop duplicates in ensembl column and keep first occurence
  mapped_ids <- mapped_ids[!duplicated(mapped_ids$ENSEMBL),]
  
  return(mapped_ids)
}

```

# Data loading and preprocessing

## Set input and output paths

```{r}
# set input and output paths
in_path <- "/mnt/Data/Projects/Cilia/revision/NonRestricted/data/"
out_path <- "/mnt/Data/Projects/Cilia/revision/NonRestricted/analysis/GO_BP/"
```

## Load data

```{r}
# Load the data
df_all <- read.delim(paste0(in_path, "All_files_combined_as_cytoscape_input.csv"), sep = "\t", header = TRUE, row.names = 1)
head(df_all)

```

## Map gene names and symbols
```{r}
# Identify columns that contain "num" in their names
num_columns <- grep("num", names(df_all), value = TRUE)

# Convert all other columns from string to logical
df_all <- df_all %>%
  mutate(across(-all_of(num_columns), ~ as.logical(.)))

# map gene IDs to gene names 
mapped_ids <- map_geneID_to_name(rownames(df_all))

# index by Ensembl
rownames(mapped_ids) <- mapped_ids$ENSEMBL

# pull out exactly one entry per row of df_all
df_all$GeneSymbol <- mapped_ids[ rownames(df_all), "SYMBOL"   ]
df_all$GeneName   <- mapped_ids[ rownames(df_all), "GENENAME" ]

# add rownames as column and reset index
df_all$Ensembl_ID <- rownames(df_all)
rownames(df_all) <- NULL

# reorder columns to have Ensembl_ID first, Symbol and Gene name first and then all other columns
df_all <- df_all %>%  dplyr::select(Ensembl_ID, GeneSymbol, GeneName, everything())
```

## Split data by location
```{r}
# Perform filtering again
df_bb <- df_all %>% filter(BasalBody_ == TRUE)
df_pc <- df_all %>% filter(PrimaryCilia_ == TRUE)
df_tip <- df_all %>% filter(PrimaryCiliaTip_ == TRUE)
df_tz <- df_all %>% filter(PrimaryCiliaTZ_ == TRUE)

# filter gene_id by location
gene_id_all <- df_all$Ensembl_ID
gene_id_bb <- df_bb$Ensembl_ID
gene_id_pc <- df_pc$Ensembl_ID
gene_id_tip <- df_tip$Ensembl_ID
gene_id_tz <- df_tz$Ensembl_ID
```

## Combine bb & tz and pc & tip
```{r}
# combine pc and tip and tz
df_pc_tip_tz <- rbind(df_pc, df_tip, df_tz)

# drop duplicates based on Ensembl_ID column
df_pc_tip_tz <- df_pc_tip_tz[!duplicated(df_pc_tip_tz$Ensembl_ID), ]
gene_id_pc_tip_tz <- df_pc_tip_tz$Ensembl_ID

```

# Compare biological themes for different locations with variable proteins

## Define variable genes

```{r}
# Load the data from the Excel file
df <- read_excel("/mnt/Data/Projects/Cilia/revision/Filtered_Staining_List_combined_exploded.xlsx")

# Identify Ensembl IDs for each category by searching substrings in the "Annotation (Intensity)" column
tip_variable <- unique(df$`Ensembl id`[grepl("Primary cilium tip", df$`Annotation (Intensity)`)])
pc_variable  <- unique(df$`Ensembl id`[grepl("Primary cilium",     df$`Annotation (Intensity)`)])
tz_variable  <- unique(df$`Ensembl id`[grepl("Primary cilium transition zone", df$`Annotation (Intensity)`)])

# Combine all variable genes and remove duplicates
variable_genes <- unique(c(pc_variable, tip_variable, tz_variable))

# Identify any non-variable genes
non_variable <- setdiff(gene_id_pc_tip_tz, variable_genes)
```

## Map as columns to df_all and save as csv file
```{r}
# add "Intensity variation" columns to df_all with true if in variable genes, false if in non_variable and NA otherwise
df_all <- df_all %>%
  mutate(
    Intensity_variation = case_when(
      Ensembl_ID %in% variable_genes ~ TRUE,
      Ensembl_ID %in% non_variable   ~ FALSE,
      TRUE                            ~ NA   # catch-all case for any other Ensembl IDs
    )
  )

# write to file
write.csv(df_all, file = paste0(out_path, "NonRestricted_all_proteins_enrichment_input_mapping.csv"))
```


## GO BP enrichment analysis
```{r}
# Split data by location
input_genes <- list(
  non_variable = non_variable,
  variable = variable_genes
)

# check length of each list
lapply(input_genes, length)

```

```{r}
comp <- compareCluster(geneCluster = input_genes,
                     fun = "enrichGO",
                     OrgDb = org.Hs.eg.db,
                     keyType       = 'ENSEMBL',
                     ont           = "BP",
                     pAdjustMethod = "BH",
                     pvalueCutoff  = 0.01,
                     qvalueCutoff  = 0.01)
```


### Dot plot of all enriched terms
```{r, fig.width=14, fig.height=50}
# plot dotplot
dotplot(comp, showCategory = NULL) + ggtitle("GO BP Enrichment Comparison", subtitle = "qvalue < 0.01") +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 100))

# save dotplot as svg file
ggsave(paste0(out_path, "NonRestricted_comparison_variable_GO_BP_dotplot.svg"), plot = last_plot(), device = "svg", width = 14, height = 50, limitsize = FALSE)

```

### Visualize overlap of enriched terms

```{r, fig.width=6, fig.height=6}
# extract results
results <- comp@compareClusterResult

# split by location
variable <- results[results$Cluster == "variable",]
non_variable <- results[results$Cluster == "non_variable",]

# Create a list of sets
go_lists <- list(
  variable = variable$ID,
  non_variable = non_variable$ID
)

# Plot the Venn diagram
venn.plot <- venn.diagram(
  x = go_lists,
  category.names = c("variable", "non_variable"),
  filename = NULL,
  output = TRUE
)

grid.newpage()
grid.draw(venn.plot)

# Save the captured plot as an SVG file
svglite(paste0(out_path, "NonRestricted_comparison_variable_GO_BP_venn.svg"), width = 6, height = 6)
grid.draw(venn.plot)
dev.off()

```

### Filter terms only enriched in one condition

```{r}

# calculate intersection of the two
unspecific_terms <- intersect(variable$ID, non_variable$ID)

# remove all unspecific terms
specific_terms <- results %>% filter(!ID %in% unspecific_terms)

# create a copy of comp
comp_filtered <- comp

# update results in comp
comp_filtered@compareClusterResult <- specific_terms
```

### Dot plot of uniquely enriched terms
```{r, fig.width=14, fig.height=50}
# plot dotplot
dotplot(comp_filtered, showCategory = NULL) + ggtitle("GO BP Enrichment Comparison", subtitle = "qvalue < 0.01")  +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 100))

# save dotplot as svg file
ggsave(paste0(out_path, "NonRestricted_comparison_variable_GO_BP_dotplot_specific_terms_only.svg"), plot = last_plot(), device = "svg", width = 14, height = 50, limitsize = FALSE)

```


```{r}
# Subset for pc_tip
variable <- comp_filtered@compareClusterResult[
  comp_filtered@compareClusterResult$Cluster == "variable", 
]

# Subset for bb_tz
non_variable <- comp_filtered@compareClusterResult[
  comp_filtered@compareClusterResult$Cluster == "non_variable", 
]

# Create new compareClusterResult objects for each subset
comp_filtered_variable <- comp_filtered
comp_filtered_non_variable <- comp_filtered

comp_filtered_variable@compareClusterResult <- variable
comp_filtered_non_variable@compareClusterResult <- non_variable

```

### Cluster results - variable proteins

```{r}
go_id = comp_filtered_variable@compareClusterResult$ID
mat = GO_similarity(go_id, 
                    ont = 'BP', 
                    db = 'org.Hs.eg.db', 
                    measure = "Sim_Relevance_2006"
                    )

```

#### Plot cluster heatmap
```{r}
# Capture the plot
heatmap_plot <- grid.grabExpr({
 df <- simplifyGO(mat,
             method = 'binary_cut',
             plot = TRUE,
             column_title = "GO BP terms only significant in variable proteins",
             use_raster = FALSE,
             order_by_size = TRUE,
             fontsize_range = c(18,36),
             max_words = 6,
             word_cloud_grob_param = list(col = 'black', 
                                          max_width = unit(200, "mm")))
})

# Save the captured plot as an SVG file
svglite(paste0(out_path, "NonRestricted_comparison_variable_GO_BP_dotplot_specific_terms_only_variable_heatmap.svg"), width = 16, height = 9)
grid.draw(heatmap_plot)
dev.off()
```

```{r, fig.width=15, fig.height=13}
grid.newpage()
grid.draw(heatmap_plot)
```

#### Process and save results

```{r}
# add cluster number from GO term clustering
results <- map_cluster_number(comp_filtered_variable,
                              df = df,
                              comp = TRUE
)

# Apply the function to each row of the DataFrame
results$GeneSymbol <- sapply(results$geneID, map_geneID_to_symbol)

# save results as csv file
write.csv(results, file = paste0(out_path, "NonRestricted_comparison_variable_GO_BP_dotplot_specific_terms_only_variable_result.csv"))
```

### Cluster results - non-variable proteins

```{r}
go_id = comp_filtered_non_variable@compareClusterResult$ID
mat = GO_similarity(go_id, 
                    ont = 'BP', 
                    db = 'org.Hs.eg.db', 
                    measure = "Sim_Relevance_2006"
                    )

```

#### Plot cluster heatmap
```{r}
# Capture the plot
heatmap_plot <- grid.grabExpr({
 df <- simplifyGO(mat,
             method = 'binary_cut',
             plot = TRUE,
             column_title = "GO BP terms only significant in non-variable proteins",
             use_raster = FALSE,
             order_by_size = TRUE,
             fontsize_range = c(18, 36),
             max_words = 6,
             word_cloud_grob_param = list(col = 'black', 
                                          max_width = unit(200, "mm")))
})

# Save the captured plot as an SVG file
svglite(paste0(out_path, "NonRestricted_comparison_variable_GO_BP_dotplot_specific_terms_only_nonvariable_heatmap.svg"), width = 16, height = 9)
grid.draw(heatmap_plot)
dev.off()
```

```{r, fig.width=16, fig.height=6}
grid.newpage()
grid.draw(heatmap_plot)
```

#### Process and save results

```{r}
# add cluster number from GO term clustering
results <- map_cluster_number(comp_filtered_non_variable,
                              df = df,
                              comp = TRUE
)

# Apply the function to each row of the DataFrame
results$GeneSymbol <- sapply(results$geneID, map_geneID_to_symbol)

# save results as csv file
write.csv(results, file = paste0(out_path, "NonRestricted_comparison_variable_GO_BP_dotplot_specific_terms_only_nonvariable_result.csv"))
```

### Filter for terms enriched in both variable and non-variable

```{r}
 # get results as data frame
# extract results
results <- comp@compareClusterResult

# split by location
variable <- results[results$Cluster == "variable",]
non_variable <- results[results$Cluster == "non_variable",]

# calculate intersection of the two
unspecific_terms <- intersect(variable$ID, non_variable$ID)

# remove all unspecific terms
specific_terms <- results %>% filter(!ID %in% unspecific_terms)
unspecific_terms <- results %>% filter(ID %in% unspecific_terms)

# create a copy of comp
comp_filtered <- comp

# update results in comp
comp_filtered@compareClusterResult <- unspecific_terms
```

### Plot dotplot of shared terms
```{r, fig.width=14, fig.height=10}
# plot dotplot
dotplot(comp_filtered, showCategory = NULL) + ggtitle("GO BP Enrichment Comparison", subtitle = "qvalue < 0.01")  +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 100))

# save dotplot as svg file
ggsave(paste0(out_path, "NonRestricted_comparison_variable_GO_BP_dotplot_shared_terms_only.svg"), plot = last_plot(), device = "svg", width = 14, height = 10, limitsize = FALSE)

```

### Cluster results - Both locations

```{r}

# Split by location
variable <- comp_filtered@compareClusterResult[comp_filtered@compareClusterResult$Cluster == "variable", ]
non_variable <- comp_filtered@compareClusterResult[comp_filtered@compareClusterResult$Cluster == "non_variable", ]

# Create new compareClusterResult objects for each subset
comp_filtered_variable<- comp_filtered
comp_filtered_non_variable <- comp_filtered

comp_filtered_variable@compareClusterResult <- variable
comp_filtered_non_variable@compareClusterResult <- non_variable

```

#### Plot cluster heatmap
```{r}
go_id = comp_filtered_variable@compareClusterResult$ID
mat = GO_similarity(go_id, 
                    ont = 'BP', 
                    db = 'org.Hs.eg.db', 
                    measure = "Sim_Relevance_2006"
                    )

```

```{r}
# Capture the plot
heatmap_plot <- grid.grabExpr({
 df <- simplifyGO(mat,
             method = 'binary_cut',
             plot = TRUE,
             column_title = "GO BP terms significant in both locations",
             use_raster = FALSE,
             order_by_size = TRUE,
             fontsize_range = c(18, 36),
             max_words = 6,
             word_cloud_grob_param = list(col = 'black', 
                                          max_width = unit(200, "mm")))
})

# Save the captured plot as an SVG file
svglite(paste0(out_path, "NonRestricted_comparison_variable_GO_BP_dotplot_shared_terms_heatmap.svg"), width = 16, height = 9)
grid.draw(heatmap_plot)
dev.off()
```

```{r, fig.width=16, fig.height=6}
grid.newpage()
grid.draw(heatmap_plot)
```

#### Process and save results

```{r}
# add cluster number from GO term clustering
results <- map_cluster_number(comp_filtered_variable,
                              df = df,
                              comp = TRUE
)

# Apply the function to each row of the DataFrame
results$GeneSymbol <- sapply(results$geneID, map_geneID_to_symbol)

# save results as csv file
write.csv(results, file = paste0(out_path, "NonRestricted_comparison_variable_GO_BP_dotplot_shared_terms_result_variable.csv"))
```

```{r}
# add cluster number from GO term clustering
results <- map_cluster_number(comp_filtered_non_variable,
                              df = df,
                              comp = TRUE
)

# Apply the function to each row of the DataFrame
results$GeneSymbol <- sapply(results$geneID, map_geneID_to_symbol)

# save results as csv file
write.csv(results, file = paste0(out_path, "NonRestricted_comparison_variable_GO_BP_dotplot_shared_terms_result_nonvariable.csv"))
```

# Session info
```{r}
sessionInfo()
```

