---
title: "Comparative functional enrichment analysis between subciliary locations for all proteins and cell types combined"
author: "Konstantin Kahnert"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    df_print: paged
    toc_float: true
    number_sections: true
    theme: flatly
    highlight: tango
  pdf_document: default
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = "CairoSVG", message = FALSE, warning = FALSE)
```

# Load libraries

```{r}
library(tidyverse)          
library(clusterProfiler)    
library(enrichplot)         
library(org.Hs.eg.db)   
library(Cairo) 
library(VennDiagram)       
library(grid)               
library(simplifyEnrichment)
```

# Define function

```{r}
# map_cluster_number: 
#   - x: enrichResult or compareClusterResult (from clusterProfiler)
#   - df: simplifyGO result data.frame with columns ID (term IDs) and Cluster_num
#   - comp: if TRUE, use x@compareClusterResult, else x@result
# Returns a data.frame of all enrich columns + term_size + Cluster_num
map_cluster_number <- function(x,
                               df,
                               comp = FALSE) {
  ## 1. Standardize the input DF’s column names
  colnames(df) <- c("ID", "Cluster_num")

  ## 2. Select the correct slot from the clusterProfiler object
  if (comp) {
    # from compareClusterResult
    results <- x@compareClusterResult
  } else {
    # from a single enrichResult
    results <- x@result
  }

  ## 3. Get term_size (the numerator of the "BgRatio" string)
  ##    e.g. "12/21273" → 12
  results$term_size <- as.numeric(
    sapply(strsplit(results$BgRatio, "/"),
           function(parts) as.numeric(parts[1]))
  )

  ## 4. Merge the enrichment results with the cluster assignments
  merged_results <- merge(
    results,
    df,
    by    = "ID",
    all.x = TRUE
  )

  ## 5. Return the processed data frame
  return(merged_results)
}


# Function to map Ensembl IDs to gene symbols for a single row
map_geneID_to_symbol <- function(geneID_str) {
  geneIDs <- unlist(strsplit(geneID_str, "/"))
  mapped_ids <- bitr(geneIDs, 
                     fromType = "ENSEMBL", 
                     toType = "SYMBOL", 
                     OrgDb = org.Hs.eg.db)
  
  # drop duplicates in ensembl column and keep first occurence
  mapped_ids <- mapped_ids[!duplicated(mapped_ids$ENSEMBL),]
  
  gene_symbols <- mapped_ids$SYMBOL
  return(paste(gene_symbols, collapse = "/"))
}


# Function to map Ensembl IDs to gene symbols for a single row
map_geneID_to_name <- function(geneIDs) {
  mapped_ids <- bitr(geneIDs, 
                     fromType = "ENSEMBL", 
                     toType = c("GENENAME", "SYMBOL"), 
                     OrgDb = org.Hs.eg.db)
  
  # drop duplicates in ensembl column and keep first occurence
  mapped_ids <- mapped_ids[!duplicated(mapped_ids$ENSEMBL),]
  
  return(mapped_ids)
}

```

# Data loading and preprocessing

## Set input and output paths

```{r}
# set input and output paths
in_path <- "/mnt/Data/Projects/Cilia/revision/NonRestricted/data/"
out_path <- "/mnt/Data/Projects/Cilia/revision/NonRestricted/analysis/GO_BP/"
```

## Load data

```{r}
# Load the data
df_all <- read.delim(paste0(in_path, "All_files_combined_as_cytoscape_input.csv"), sep = "\t", header = TRUE, row.names = 1)
head(df_all)

```

## Map gene names and symbols
```{r}
# Identify columns that contain "num" in their names
num_columns <- grep("num", names(df_all), value = TRUE)

# Convert all other columns from string to logical
df_all <- df_all %>%
  mutate(across(-all_of(num_columns), ~ as.logical(.)))

# map gene IDs to gene names 
mapped_ids <- map_geneID_to_name(rownames(df_all))

# index by Ensembl
rownames(mapped_ids) <- mapped_ids$ENSEMBL

# pull out exactly one entry per row of df_all
df_all$GeneSymbol <- mapped_ids[ rownames(df_all), "SYMBOL"   ]
df_all$GeneName   <- mapped_ids[ rownames(df_all), "GENENAME" ]

# add rownames as column and reset index
df_all$Ensembl_ID <- rownames(df_all)
rownames(df_all) <- NULL

# reorder columns to have Ensembl_ID first, Symbol and Gene name first and then all other columns
df_all <- df_all %>%  dplyr::select(Ensembl_ID, GeneSymbol, GeneName, everything())
```

## Split data by location
```{r}
# Perform filtering again
df_bb <- df_all %>% filter(BasalBody_ == TRUE)
df_pc <- df_all %>% filter(PrimaryCilia_ == TRUE)
df_tip <- df_all %>% filter(PrimaryCiliaTip_ == TRUE)
df_tz <- df_all %>% filter(PrimaryCiliaTZ_ == TRUE)
df_nucleus <- df_all %>% filter(Nucleus == TRUE)
df_mitotic <- df_all %>% filter(Mitotic == TRUE)
df_membrane <- df_all %>% filter(Membrane == TRUE)
df_cytoplasm <- df_all %>% filter(Cytoplasm == TRUE)

# filter gene_id by location
gene_id_all <- df_all$Ensembl_ID
gene_id_bb <- df_bb$Ensembl_ID
gene_id_pc <- df_pc$Ensembl_ID
gene_id_tip <- df_tip$Ensembl_ID
gene_id_tz <- df_tz$Ensembl_ID
gene_id_nucleus <- df_nucleus$Ensembl_ID
gene_id_mitotic <- df_mitotic$Ensembl_ID
gene_id_membrane <- df_membrane$Ensembl_ID
gene_id_cytoplasm <- df_cytoplasm$Ensembl_ID

```

## Combine bb & tz and pc & tip
```{r}
# combine bb and tz
df_bb_tz <- rbind(df_bb, df_tz)

# drop duplicates based on Ensembl_ID column
df_bb_tz <- df_bb_tz[!duplicated(df_bb_tz$Ensembl_ID), ]
gene_id_bb_tz <- df_bb_tz$Ensembl_ID


# combine pc and tip
df_pc_tip <- rbind(df_pc, df_tip)

# drop duplicates based on Ensembl_ID column
df_pc_tip <- df_pc_tip[!duplicated(df_pc_tip$Ensembl_ID), ]
gene_id_pc_tip <- df_pc_tip$Ensembl_ID


# combine pc and tip and tz
df_pc_tip_tz <- rbind(df_pc, df_tip, df_tz)

# drop duplicates based on Ensembl_ID column
df_pc_tip_tz <- df_pc_tip_tz[!duplicated(df_pc_tip_tz$Ensembl_ID), ]
gene_id_pc_tip_tz <- df_pc_tip_tz$Ensembl_ID

# get intersection of all genes
gene_id_all_cilia <- Reduce(union, list(gene_id_bb_tz, gene_id_pc_tip))
print(paste("Number of genes in all cilia locations:", length(gene_id_all_cilia)))
```

# All proteins

## GO BP enrichment analysis

```{r}
ego <- enrichGO(gene          = gene_id_all,
                OrgDb         = org.Hs.eg.db,
                keyType       = 'ENSEMBL',
                ont           = "BP",
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.01,
                qvalueCutoff  = 0.01,
                readable      = TRUE)

ego@result <- ego@result[ego@result$p.adjust < 0.01,]
ego

```


### Dot plot of top 100 enriched proteins

```{r, dev = 'svg', fig.width=10, fig.height=20}
dotplot(ego, showCategory=100, label_format = 50) + ggtitle("All Cilia Proteins", subtitle = "Top 100 significantly overrepresented GO BP terms") +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 70))

# save dotplot as svg file
ggsave(paste0(out_path, "NonRestricted_all_proteins_enrich_GO_BP_dotplot.svg"), plot = last_plot(), device = "svg", width = 10, height = 20, limitsize = FALSE)

```

# Compare biological themes for different locations (cilia only, PC+tip vs TZ+BB)

```{r}
# prepare input list
input_genes <- list(
    PC_tip = gene_id_pc_tip,
    BB_TZ = gene_id_bb_tz
)
```

## GO BP enrichment analysis

```{r}
# Perform the compareCluster analysis
comp <- compareCluster(geneCluster = input_genes,
                     fun = "enrichGO",
                     OrgDb = org.Hs.eg.db,
                     keyType       = 'ENSEMBL',
                     ont           = "BP",
                     pAdjustMethod = "BH",
                     pvalueCutoff  = 0.01,
                     qvalueCutoff  = 0.01)
```

### Dot plot of all enriched terms
```{r, dev = 'svg', fig.width=14, fig.height=70, eval=TRUE}
dotplot(comp, showCategory = NULL) + ggtitle("GO BP Enrichment Comparison", subtitle = "qvalue < 0.01") +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 100))

# save dotplot as svg file
ggsave(paste0(out_path, "NonRestricted_comparison_locations_cilia_only_combined_GO_BP_dotplot.svg"), plot = last_plot(), device = "svg", width = 14, height = 40)

```

### Visualize overlap of enriched terms as Venn diagram

```{r, dev = 'svg', fig.width=7, fig.height=6}
# extract results
results <- comp@compareClusterResult

# split by location
pc_tip <- results[results$Cluster == "PC_tip",]
bb_tz <- results[results$Cluster == "BB_TZ",]

# Create a list of the four sets
go_lists <- list(
  PC_tip = pc_tip$ID,
  BB_TZ = bb_tz$ID
)

# Plot the Venn diagram
venn.plot <- venn.diagram(
  x = go_lists,
  category.names = c("PC+tip", "BB+TZ"),
  filename = NULL,
  output = TRUE
)

grid.newpage()
grid.draw(venn.plot)

# Save the captured plot as an SVG file
svg(paste0(out_path, "NonRestricted_comparison_locations_cilia_only_combined_GO_BP_venn.svg"), width = 7, height = 6)
grid.draw(venn.plot)
dev.off()

```

### Determine uniquely enriched and unspecifically enriched terms

```{r}
# calculate intersection of the two
unspecific_terms <- intersect(pc_tip$ID, bb_tz$ID)

# remove all unspecific terms
specific_terms <- results %>% filter(!ID %in% unspecific_terms)

# create a copy of comp
comp_filtered <- comp

# update results in comp
comp_filtered@compareClusterResult <- specific_terms
```

### Dot plot of uniquely enriched terms
```{r, dev = 'svg', fig.width=14, fig.height=50}
# plot(xx, type="dot", caption="GO Enrichment Comparison")
dotplot(comp_filtered, showCategory = NULL) + ggtitle("GO BP Enrichment Comparison", subtitle = "qvalue < 0.01")  +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 100))

# save dotplot as svg file
ggsave(paste0(out_path, "NonRestricted_comparison_locations_cilia_only_combined_GO_BP_dotplot_specific_terms_only.svg"), plot = last_plot(), device = "svg", width = 14, height = 50, limitsize = FALSE)

```

## Prepare data for clustering of enriched terms

```{r}

# Split by location
pc_tip <- comp_filtered@compareClusterResult[comp_filtered@compareClusterResult$Cluster == "PC_tip", ]
bb_tz <- comp_filtered@compareClusterResult[comp_filtered@compareClusterResult$Cluster == "BB_TZ", ]

# Subset for pc_tip
pc_tip <- comp_filtered@compareClusterResult[
  comp_filtered@compareClusterResult$Cluster == "PC_tip", 
]

# Subset for bb_tz
bb_tz <- comp_filtered@compareClusterResult[
  comp_filtered@compareClusterResult$Cluster == "BB_TZ", 
]

# Create new compareClusterResult objects for each subset
comp_filtered_pc_tip <- comp_filtered
comp_filtered_bb_tz <- comp_filtered

comp_filtered_pc_tip@compareClusterResult <- pc_tip
comp_filtered_bb_tz@compareClusterResult <- bb_tz

```

## Cluster results - PC&tip

```{r}
go_id = pc_tip$ID
mat = GO_similarity(go_id, 
                    ont = 'BP', 
                    db = 'org.Hs.eg.db', 
                    measure = "Sim_Relevance_2006"
                    )

```


### Plot cluster heatmap 
```{r}
# Capture the plot
heatmap_plot <- grid.grabExpr({
 df <- simplifyGO(mat,
             method = 'binary_cut',
             plot = TRUE,
             column_title = "GO BP terms only significant in primary cilia incl. tip",
             use_raster = FALSE,
             order_by_size = TRUE,
             fontsize_range = c(9, 18),
             max_words = 6,
             word_cloud_grob_param = list(col = 'black', 
                                          max_width = unit(200, "mm")))
})

# Save the captured plot as an SVG file
svg(paste0(out_path, "NonRestricted_comparison_locations_cilia_only_combined_GO_BP_dotplot_specific_terms_only_PC_heatmap.svg"), width = 16, height = 8)
grid.draw(heatmap_plot)
dev.off()
```

```{r, dev = 'svg', fig.width=15, fig.height=13}
grid.newpage()
grid.draw(heatmap_plot)
```

### Process and save results

```{r}
# add cluster number from GO term clustering
results <- map_cluster_number(comp_filtered_pc_tip,
                              df = df,
                              comp = TRUE
)

# Apply the function to each row of the DataFrame
results$GeneSymbol <- sapply(results$geneID, map_geneID_to_symbol)

# save results as csv file
write.csv(results, file = paste0(out_path, "NonRestricted_comparison_locations_cilia_only_combined_GO_BP_dotplot_specific_terms_only_PC_result.csv"))
```


## Cluster results - BB&TZ

```{r}
go_id = bb_tz$ID
mat = GO_similarity(go_id, 
                    ont = 'BP', 
                    db = 'org.Hs.eg.db', 
                    measure = "Sim_Relevance_2006"
                    )

```

### Plot cluster heatmap 
```{r}
# Capture the plot
heatmap_plot <- grid.grabExpr({
 df <- simplifyGO(mat,
             method = 'binary_cut',
             plot = TRUE,
             column_title = "GO BP terms only significant in BB & TZ",
             use_raster = FALSE,
             order_by_size = TRUE,
             fontsize_range = c(9, 18),
             max_words = 6,
             word_cloud_grob_param = list(col = 'black', 
                                          max_width = unit(200, "mm")))
})

# Save the captured plot as an SVG file
svg(paste0(out_path, "NonRestricted_comparison_locations_cilia_only_combined_GO_BP_dotplot_specific_terms_only_BB_heatmap.svg"), width = 16, height = 8)
grid.draw(heatmap_plot)
dev.off()
```

```{r, dev = 'svg', fig.width=15, fig.height=13}
grid.newpage()
grid.draw(heatmap_plot)
```

### Process and save results

```{r}
# add cluster number from GO term clustering
results <- map_cluster_number(comp_filtered_bb_tz,
                              df = df,
                              comp = TRUE
)

# Apply the function to each row of the DataFrame
results$GeneSymbol <- sapply(results$geneID, map_geneID_to_symbol)

# save results as csv file
write.csv(results, file = paste0(out_path, "NonRestricted_comparison_locations_cilia_only_combined_GO_BP_dotplot_specific_terms_only_BB_result.csv"))
```

## Filter for terms enriched in both locations

```{r}
 # get results as data frame
comp_results <- comp@compareClusterResult

# get all terms significant in each location
pc_tip <- comp_results %>% filter(Cluster == "PC_tip")
bb_tz <- comp_results %>% filter(Cluster == "BB_TZ")

# calculate intersection of the two
unspecific_terms <- intersect(pc_tip$ID, bb_tz$ID)

# remove all unspecific terms
unspecific_terms <- comp_results %>% filter(ID %in% unspecific_terms)

# create a copy of comp
comp_filtered <- comp

# update results in comp
comp_filtered@compareClusterResult <- unspecific_terms
```

## Dot plot of shared terms
```{r, dev = 'svg', fig.width=14, fig.height=50}
# plot dotplot
dotplot(comp_filtered, showCategory = NULL) + ggtitle("GO BP Enrichment Comparison", subtitle = "qvalue < 0.01")  +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 100))

# save dotplot as svg file
ggsave(paste0(out_path, "NonRestricted_comparison_locations_cilia_only_combined_GO_BP_dotplot_shared_terms_only.svg"), plot = last_plot(), device = "svg", width = 14, height = 50, limitsize = FALSE)

```


## Prepare data for clustering of shared enriched terms
```{r}
# Split by location
pc_tip <- comp_filtered@compareClusterResult[comp_filtered@compareClusterResult$Cluster == "PC_tip", ]
bb_tz <- comp_filtered@compareClusterResult[comp_filtered@compareClusterResult$Cluster == "BB_TZ", ]

# Create new compareClusterResult objects for each subset
comp_filtered_pc_tip <- comp_filtered
comp_filtered_bb_tz <- comp_filtered

comp_filtered_pc_tip@compareClusterResult <- pc_tip
comp_filtered_bb_tz@compareClusterResult <- bb_tz

```

## Cluster results - Both locations
```{r}
go_id = pc_tip$ID
mat = GO_similarity(go_id, 
                    ont = 'BP', 
                    db = 'org.Hs.eg.db', 
                    measure = "Sim_Relevance_2006"
                    )

```

### Plot cluster heatmap 
```{r}
# Capture the plot
heatmap_plot <- grid.grabExpr({
 df <- simplifyGO(mat,
             method = 'binary_cut',
             plot = TRUE,
             column_title = "GO BP terms significant in both locations",
             use_raster = FALSE,
             order_by_size = TRUE,
             fontsize_range = c(9, 18),
             max_words = 6,
             word_cloud_grob_param = list(col = 'black', 
                                          max_width = unit(200, "mm")))
})

# Save the captured plot as an SVG file
svg(paste0(out_path, "NonRestricted_comparison_locations_cilia_only_combined_GO_BP_dotplot_specific_terms_BB&PC_heatmap.svg"), width = 16, height = 8)
grid.draw(heatmap_plot)
dev.off()
```

```{r, dev = 'svg', fig.width=15, fig.height=13}
grid.newpage()
grid.draw(heatmap_plot)
```

### Process and save results

```{r}
# add cluster number from GO term clustering
results <- map_cluster_number(comp_filtered_pc_tip,
                              df = df,
                              comp = TRUE
)

# Apply the function to each row of the DataFrame
results$GeneSymbol <- sapply(results$geneID, map_geneID_to_symbol)

# save results as csv file
write.csv(results, file = paste0(out_path, "NonRestricted_comparison_locations_cilia_only_combined_GO_BP_dotplot_specific_terms_BB&PC_result_PC.csv"))
```

```{r}
# add cluster number from GO term clustering
results <- map_cluster_number(comp_filtered_bb_tz,
                              df = df,
                              comp = TRUE
)

# Apply the function to each row of the DataFrame
results$GeneSymbol <- sapply(results$geneID, map_geneID_to_symbol)

# save results as csv file
write.csv(results, file = paste0(out_path, "NonRestricted_comparison_locations_cilia_only_combined_GO_BP_dotplot_specific_terms_BB&PC_result_BB.csv"))

```

# Compare biological themes for different locations (cilia vs other parts of cell)

```{r}
# Split data by location
input_genes <- list(
  PrimaryCiliaTip = gene_id_tip,
  PrimaryCilia = gene_id_pc,
  PrimaryCiliaTZ = gene_id_tz,
  BasalBody = gene_id_bb,
  Cytoplasm = gene_id_cytoplasm,
  Membrane = gene_id_membrane,
  Nucleus = gene_id_nucleus,
  Mitotic = gene_id_mitotic
)
```

## GO BP enrichment analysis

```{r}

# Perform the compareCluster analysis
comp <- compareCluster(geneCluster = input_genes,
                     fun = "enrichGO",
                     OrgDb = org.Hs.eg.db,
                     keyType       = 'ENSEMBL',
                     ont           = "BP",
                     pAdjustMethod = "BH",
                     pvalueCutoff  = 0.01,
                     qvalueCutoff  = 0.01)
```

```{r}
# get results 
results <- comp@compareClusterResult

# Apply the function to each row of the DataFrame
results$GeneSymbol <- sapply(results$geneID, map_geneID_to_symbol)

# save results as csv file
write.csv(results, file = paste0(out_path, "NonRestricted_comparison_all_locations_GO_BP_result.csv"))
```

### Dot plot of all enriched terms
```{r, dev = 'svg', fig.width=14, fig.height=120, eval=TRUE}
dotplot(comp, showCategory = NULL) + ggtitle("GO BP Enrichment Comparison", subtitle = "qvalue < 0.01") +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 60))

# save dotplot as svg file
ggsave(paste0(out_path, "NonRestricted_comparison_all_locations_GO_BP_dotplot.svg"), plot = last_plot(), device = "svg", width = 16, height = 350, limitsize = FALSE)

```


```{r}
# extract results
results <- comp@compareClusterResult$ID

# get unique terms
results <- unique(results)

# filter for p.adjust < 0.01 and qvalue < 0.01
results_signif <- comp@compareClusterResult %>% filter(p.adjust < 0.01 & qvalue < 0.01)
results_signif <- unique(results_signif$ID)

# print number of unique terms
print(paste("Number of unique terms in comp:", length(results)))
print(paste("Number of unique terms in comp:", length(results_signif)))

```

### Filter for cilia terms

Here we filter for terms that are enriched in any of the four cilia locations, so basically cutting of the long dotplot above to only show the terms of the cilia locations.

```{r}
 # get results as data frame
comp_results <- comp@compareClusterResult

# get all terms that are significant in any of the 4 non cilia locations
non_cilia_terms <- comp_results %>% filter(Cluster == "Cytoplasm" | Cluster == "Membrane" | Cluster == "Nucleus" | Cluster == "Mitotic") 

# get all the terms that are significant in any of the cilia locations
cilia_terms <- comp_results %>% filter(Cluster == "PrimaryCilia" | Cluster == "PrimaryCiliaTip" | Cluster == "PrimaryCiliaTZ" | Cluster == "BasalBody")

# remove all cilia terms from non_cilia_terms
non_cilia_only_terms <- non_cilia_terms %>% filter(!ID %in% cilia_terms$ID)

# remove all non cilia only terms from cilia terms
cilia_only_terms <- cilia_terms %>% filter(!ID %in% non_cilia_only_terms$ID)

# filter comp results for terms in cilia_only_terms
comp_results_filtered <- comp_results %>% filter(ID %in% cilia_only_terms$ID)

# create a copy of comp
comp_filtered <- comp

# update results in comp
comp_filtered@compareClusterResult <- comp_results_filtered
```

### Dot plot of terms enriched in ciliary locations
```{r, dev = 'svg', fig.width=16, fig.height=100, eval=TRUE}
# plot dotplot
dotplot(comp_filtered, showCategory = NULL) + ggtitle("GO BP Enrichment Comparison", subtitle = "qvalue < 0.01") +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 80))

# save dotplot as svg file
ggsave(paste0(out_path, "NonRestricted_comparison_all_comparison_locations_GO_BP_dotplot_filtered_cilia_terms.svg"), plot = last_plot(), device = "svg", width = 16, height = 100, limitsize = FALSE)

```

### Filter for cilia only terms

Here we filter for terms that are only enriched in any of the four cilia locations and not in any of the other four non cilia locations. Note: This removes the non cilia locations from the plot as there is no enrichment of any of the terms in any non cilia location.

```{r}
 # get results as data frame
comp_results <- comp@compareClusterResult

# get all terms that are significant in any of the 4 non cilia locations
non_cilia_terms <- comp_results %>% filter(Cluster == "Cytoplasm" | Cluster == "Membrane" | Cluster == "Nucleus" | Cluster == "Mitotic") 

# remove all non cilia terms from results
cilia_only_terms <- comp_results %>% filter(!ID %in% non_cilia_terms$ID)

# create a copy of comp
comp_filtered <- comp

# update results in comp
comp_filtered@compareClusterResult <- cilia_only_terms
```

### Dot plot of terms only enriched in ciliary locations
```{r, dev = 'svg', fig.width=14, fig.height=30}
# plot dotplot
dotplot(comp_filtered, showCategory = NULL) + ggtitle("GO BP Enrichment Comparison", subtitle = "qvalue < 0.01") +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 100))

# save dotplot as svg file
ggsave(paste0(out_path, "NonRestricted_comparison_all_comparison_locations_GO_BP_dotplot_filtered_cilia_only_terms.svg"), plot = last_plot(), device = "svg", width = 16, height = 30, limitsize = FALSE)

```

```{r}
# extract results
results <- comp@compareClusterResult$ID

# get unique terms
results <- unique(results)

# print number of unique terms
print(paste("Number of unique terms in comp:", length(results)))
```

### Simplify results

```{r}
comp_simplified <- simplify(comp, cutoff=0.6, by="p.adjust", select_fun=min)

# print number of unique terms
print(paste("Number of signficant terms:", nrow(comp)))
print(paste("Number of significant terms after simplifying:", nrow(comp_simplified)))

```

```{r}
# extract results
results <- comp_simplified@compareClusterResult$ID

# get unique terms
results <- unique(results)

# print number of unique terms
print(paste("Number of unique terms in comp_simplified:", length(results)))
```

```{r}
# get results 
results <- comp_simplified@compareClusterResult

# Apply the function to each row of the DataFrame
results$GeneSymbol <- sapply(results$geneID, map_geneID_to_symbol)

# save results as csv file
write.csv(results, file = paste0(out_path, "NonRestricted_comparison_all_locations_GO_BP_result_simplified.csv"))
```

### Dot plot of all enriched terms (simplified)
```{r, dev = 'svg', fig.width=14, fig.height=50}
# plot dotplot
dotplot(comp_simplified, showCategory = NULL) + ggtitle("GO BP Enrichment Comparison", subtitle = "qvalue < 0.01") +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 80))
```

### Filter for cilia terms

Here we filter for terms that are enriched in any of the four cilia locations, so basically cutting of the long dotplot above to only show the terms of the cilia locations. But this time using the simplified results.

```{r}
 # get results as data frame
comp_results <- comp_simplified@compareClusterResult

# get all terms that are significant in any of the 4 non cilia locations
non_cilia_terms <- comp_results %>% filter(Cluster == "Cytoplasm" | Cluster == "Membrane" | Cluster == "Nucleus" | Cluster == "Mitotic") 

# get all the terms that are significant in any of the cilia locations
cilia_terms <- comp_results %>% filter(Cluster == "PrimaryCilia" | Cluster == "PrimaryCiliaTip" | Cluster == "PrimaryCiliaTZ" | Cluster == "BasalBody")

# remove all cilia terms from non_cilia_terms
non_cilia_only_terms <- non_cilia_terms %>% filter(!ID %in% cilia_terms$ID)

# remove all non cilia only terms from cilia terms
cilia_only_terms <- cilia_terms %>% filter(!ID %in% non_cilia_only_terms$ID)

# filter comp results for terms in cilia_only_terms
comp_results_filtered <- comp_results %>% filter(ID %in% cilia_only_terms$ID)

# create a copy of comp
comp_simplified_filtered <- comp_simplified

# update results in comp
comp_simplified_filtered@compareClusterResult <- comp_results_filtered
```

### Dot plot of terms enriched in ciliary locations (simplified)
```{r, dev = 'svg', fig.width=14, fig.height=40}
# plot dotplot
dotplot(comp_simplified_filtered, showCategory = NULL) + ggtitle("GO BP Enrichment Comparison", subtitle = "qvalue < 0.01") +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 80))

# save dotplot as svg file
ggsave(paste0(out_path, "NonRestricted_comparison_all_comparison_locations_GO_BP_dotplot_simplified_06_cilia_terms.svg"), plot = last_plot(), device = "svg", width = 16, height = 40, limitsize = FALSE)

```

### Filter for cilia only terms

Here we filter again for terms that are only enriched in any of the four cilia locations and not in any of the other four non cilia locations. But this time using the simplified results. Note: This removes the non cilia locations from the plot as there is no enrichment of any of the terms in any non cilia location.

```{r}
 # get results as data frame
comp_results <- comp_simplified@compareClusterResult

# get all terms that are significant in any of the 4 non cilia locations
non_cilia_terms <- comp_results %>% filter(Cluster == "Cytoplasm" | Cluster == "Membrane" | Cluster == "Nucleus" | Cluster == "Mitotic") 

# remove all non cilia terms from results
cilia_only_terms <- comp_results %>% filter(!ID %in% non_cilia_terms$ID)

# create a copy of comp
comp_simplified_filtered <- comp_simplified

# update results in comp
comp_simplified_filtered@compareClusterResult <- cilia_only_terms
```

### Dot plot of terms only enriched in ciliary locations(simplified)
```{r, dev = 'svg', fig.width=14, fig.height=20}
# plot dotplot
dotplot(comp_simplified_filtered, showCategory = NULL) + ggtitle("GO BP Enrichment Comparison", subtitle = "qvalue < 0.01") +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 80))

# save dotplot as svg file
ggsave(paste0(out_path, "NonRestricted_comparison_all_comparison_locations_GO_BP_dotplot_simplified_06_cilia_only_terms.svg"), plot = last_plot(), device = "svg", width = 16, height = 20, limitsize = FALSE)

```

#### Check overlap between significant terms and simplified terms

```{r, dev = 'svg', fig.width=5, fig.height=5}
# calculate overlap between comp and comp_simplified results by plotting a venn diagramm
# extract results
results <- comp@compareClusterResult$ID
results_simplified <- comp_simplified@compareClusterResult$ID


# print number of unique terms
print(paste("Number of terms in comp:", length(results)))
print(paste("Number of terms in comp_simplified:", length(results_simplified)))

# get unique terms
results <- unique(results)
results_simplified <- unique(results_simplified)

# print number of unique terms
print(paste("Number of unique terms in comp:", length(results)))
print(paste("Number of unique terms in comp_simplified:", length(results_simplified)))

intersection <- intersect(results, results_simplified)
print(paste("Number of terms in intersection:", length(intersection)))

# plot venn diagram
venn.plot <- venn.diagram(
  x = list(results = results, results_simplified = results_simplified),
  category.names = c("All terms", "Simplified terms"),
  filename = NULL,
  output = FALSE
)

grid.newpage()
grid.draw(venn.plot)

```

# Session info
```{r}
sessionInfo()
```

